---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  collapse = TRUE,
  comment = "#>"
)
```

# pd1-enhdel-omics

<!-- badges: start -->
<!-- badges: end -->

This repository provides code to reproduce results and figures from **Deletion of an exhaustion-specific PD-1 enhancer improves T cell function and viral control**.

The associated GEO repository is GSE212507. Additionally, the single cell counts matrices and other associated data tables have been uploaded to the Single Cell Portal under the [PD-1 enhancer deletion](https://singlecell.broadinstitute.org/single_cell/study/SCP1772/pd-1-enhancer-deletion). 

## Contents

 - `scRNAseq`: directory containing the data and scripts for reproducing scRNA-seq analyses and figures
 - `ATACseq`: directory containing the data and scripts for reproducing ATAC-seq analyses and figures
 - `figures`: directory containing the scripts to reproduce the figures
 - `renv`: directory containing all information for tracking R package versioning and downloads. See [Introduction to renv](https://rstudio.github.io/renv/articles/renv.html) for details
 - `aesthetics.R`: file containing aesthetics and themes for the project
 - `helper.R`: file containing libraries and helper functions
 - `driver.R`: optional script to regenerate all intermediate results. May take a long time (> 24 hours) to execute. 
 
## Instructions

To regenerate figures from raw data, take the following steps: 

1. Clone the GitHub. 
2. Restore the renv environment from the renv.lock file, using `renv::restore()`. The [VISION package](https://github.com/YosefLab/VISION), a signature scoring method developed by the Yosef lab, may have to be manually installed from their GitHub repository. The package can be installed using `devtools` by running the code below: 

```
library(devtools)
devtools::install_github("YosefLab/VISION")
```

3. To reproduce the footprinting analysis, you must also install an appropriate conda environment to run the TOBIAS command line tool. You can do this by [creating a conda environment from the YML file](https://docs.conda.io/projects/conda/en/latest/user-guide/tasks/manage-environments.html#creating-an-environment-from-an-environment-yml-file) `tobias3.7_env.yml`. 
3. Download the single cell counts matrices from the Single Cell Portal and place them in the `scRNAseq/data/counts_matrices` directory. 
4. For the footprinting analysis, download the ATACseq bam files from the Single Cell Portal and place them in the `ATACseq/data/bams` directory. Also download the compressed directory `footprinting.zip` and place it at `ATACseq/data/footprinting`.
5. Regenerate all intermediate data tables and objects by running the script `driver.R` and `footprinting.sh`. Warning: It may take a long time, >24 hours depending on compute power, to execute. You can blame the signature scoring script for this; other steps are relatively quick. 
6. Use Rmarkdown to knit the `.Rmd` files in the `figures/` directory to output the figures in either HTML or PDF format. 

Alternatively, to bypass the long compute time in step 3, follow the instructions below to generate/acquire processed data files, then knit the `.Rmd` files in the `figures/` directory. 

1. Run the script at `scRNAseq/scripts/02_gene_signatures` to generate the `scRNAseq/data/gene_signatures/02_gene_signatures.msigdbr_H-C2-C7.tsv`
2. Download additional processed data files from the Study Files section [Single Cell Portal](https://singlecell.broadinstitute.org/single_cell/study/SCP1772/pd-1-enhancer-deletion#study-download) and place them at the following paths. 

* scRNAseq/data/counts_matrices.zip -> decompress
* scRNAseq/data/processed_data_tables.zip -> decompress
* scRNAseq/data/processed_data_objects/00_seurat_workflow.so.rds
* scRNAseq/data/processed_data_objects/05_pseudobulk.deseq_list.rds
* ATACseq/data/processed_data_tables/01_analyze.deseq_results_tidy.tsv

## Contact 

Please contact a corresponding author for any questions, comments, or concerns regarding the paper in general.

For issues with code/reproducibility, please open a GitHub issue. 

## Compute environment and version

```{r}
sessionInfo()
```

